A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-07-07 demystify.tex
updates

For external fellowships including those from the US government (e.g., NSF, DOD) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.  % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
Prestigious external fellowships typically require a clear and good research plan, so it might be a good idea to wait until at least your second year to have research experience and even publication before applying. Remember, you're competing with the top PhD students at top universities worldwide.

@@

In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.  Every PhD student has pubs, but only a few would have NSF Graduate (GRFP) or Microsoft fellowships. In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, you school will still mention you on their website, and you should put it on your CV.

@@

\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around around \textbf{April 15th} in the US. This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more). After this date, CS programs can gauge how many slots remain unfilled.

>>>>

For external fellowships including those from the US government (e.g., NSF, DOD) and tech companies (e.g., Google, Microsoft, Facebook), you will need to apply.  % However, tech companies including Google, Microsoft, Facebook have fellowships that international students can apply for.
Prestigious external fellowships typically require a clear and good research plan (the GRFP also requires  outreach activities and plans). So it might be a good idea to wait until your second year to have research experience and even publication before applying. Remember, you're competing with the top PhD students at top universities worldwide.

@@

In general, fellowships, especially those that are open to everyone, are highly competitive and prestigious, and you will stand out if you get one.  Every PhD student has pubs, but only a few would have NSF Graduate (GRFP\footnote{\url{https://www.alexhunterlang.com/nsf-fellowship} is a good starting place for the GRFP with lots of proposal examples.}) or Microsoft fellowships. In fact, these are so prestigious that even if you didn't get it but make it to the final round or even \emph{``honorable mentioning''}, your school will still mention you on their website, and you should put it on your CV.

@@

\paragraph{Response Deadlines} Accepted students are usually given a deadline to make decision on their offers, often around around \textbf{April 15th} in the US. This is a standard deadline for PhD programs due to the Council of Graduate Schools (CGS) Resolution (\href{https://cgsnet.org/wp-content/uploads/2024/01/CGS_April15_Resolution_Jan312024.pdf}{see more}. After this date, CS programs can gauge how many slots remain unfilled.

### 2024-07-07 faculty.tex
updates

While only 10\% of your work is officially for services, they are very important for your tenure and career.
%You are expected to have certain types of services as a faculty member.  
Some services are required (e.g., when your department chair assigns you to some committee), but some are not and you can decline (e.g., an invitation to be a PC member).  You should be strategic about which services you accept.

\section{Service to your Research Community}

These are services that you do for your research community. They will help you to get to know people in your field, and to get your name out.  Common examples include serving on the program committees of conferences, serving on editorial boards of journal, and serving on grant panels.

@@

\subsection{Volunteer for more Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.  For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.  These roles are more time-consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.

@@

\subsection{Personal Experience} In my case I mostly cold email people for opportunities. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request (I did have papers at OOPSLA and PLDI when I asked). I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE. I also got invited to serve on a PC because a senior colleague recommended me.
For NSF, I emailed program managers and asked to serve on their panels.

I felt that I didn't get much out of serving on PCs and didn't really like the review process much. However, I still do it because I feel it is my responsibility give back to the community.

@@

I have also volunteered for leadership roles, e.g., I directly asked the general chair of ICSE'21 to see if I can take more leadership roles and was then invited to serve as a proceeding co-chair. In hindsight this particular role is a lot of work and I didn't really like many aspects of it, but it was an eye-openning experience.

\section{Service to the Department}
Your department chair will assign you to some committees to help run the department.  Typically for tenure track faculty the services are lighter than tenured ones (e.g., less leadership roles). Common committees include PhD and/or MS admission, faculty search, and colloquium committees.

Depends on your preference and how you are perceived, you can take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.  These roles are more time consuming and might involve some politics, but they are also more rewarding and can help your tenure case.

@@

A good department will try to match you with committees that can help junior faculty build their lab and tenure case (and shield you from committees that are not as useful).

\subsection{Personal Experience} Since coming to GMU, I have been very active in serving the department.  I share the vision of the department and want to help achieve it. In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was my study break) and help recruit strong ones. I also agree to be director of the MS SWE program, which is a lot of work and typically done by tenured faculty.

Occasionally I feel overwhelmed by having many services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues (and also head butt with some), and that is important for my tenure case.
In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to run a colloquium, etc).

In short, if you want to make a difference, and \emph{feel comfortable} at handling additional load, you can volunteer for more services. Otherwise, you can just ``stay low'',
do the required services and focus on your research and teaching.

\section{Service to the University}
\section{Other Types of Services}
\paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and colleagues.

>>>>

While only 10\% of your work is officially for services, they are very important for your tenure and career.
%You are expected to have certain types of services as a faculty member.  
Some services are required (e.g., when your department chair assigns you to some committee), but some are not and you can decline (e.g., an invitation to be a PC member).

\section{Service to Research Community}

These are services that you do for your research community. They will help you to get to know people in your field, and to get your name out.  Common examples include serving on the program committees (PC) of conferences, serving on editorial boards of journal, and serving on grant panels (e.g., from NSF).

@@

\subsection{Volunteer for Leadership roles} Once you have served on a few PCs and NSF panels, you should aim for more leadership roles.  For example, you can volunteer to be a chair of certain sessions/tracks of a large conference, or even a general chair of a workshop or conference.  These roles are more time-consuming, but they can be more rewarding (e.g., more people will know about you) and in general can help your career, especially your tenure case.

@@

\subsection{Personal Experience} In my case I mostly cold email people for opportunities. For example, I directly email OOPSLA and PLDI PC chairs and asked, and they accepted my request. I tried this with SE conferences  e.g., ICSE (and also nominate myself through the official form), and it never works, even though I have many papers at ICSE. I also got invited to serve on a PC because a senior colleague recommended me.
For NSF, I emailed program managers and asked to serve on their panels.

I felt that I didn't get much out of serving on PCs and didn't really like the review process much. However, I still do it because I feel it is my responsibility, i.e., people spent time reviewing my papers, so I should also volunteer my time to review theirs.

@@

I have also volunteered for leadership roles, e.g., I directly asked the general chair of ICSE'21 to see if I can take more leadership roles and was then invited to serve as a proceeding co-chair. In hindsight this particular role is a lot of work, and I didn't really like many aspects of it, but it was an eye-openning experience.

\section{Service to the Department}
Your department chair will assign you to some committees to help run the department.  For tenure-track faculty the services are often lighter than tenured ones (e.g., less leadership roles). Common committees include PhD and/or MS admission, faculty search, and colloquium committees.

You can also take more leadership roles when you are close to tenure, such as being the chair of a committee or chair/director of a program.  These are more time-consuming and might involve some politics, but they are also more rewarding and can help your tenure case.

@@

A good department/chair will try to match you with committees that can help junior faculty build their lab and tenure case (and shield you from committees that are not as useful).

\subsection{Personal Experience} Since coming to GMU, I have been very active in serving the department.  In addition to the typical services load, I volunteer to serve in additional committees and leadership roles. For example, I maintain the CSRankings DB for the department, organize the Open House to recruit admitted students, participate in the Dept Chair renewal committee, etc.
Even though I am not on the search committee, I meet with every faculty candidate (I talked to 22 people in Spring'23, which was during my study break). I also agree to be director of the MS SWE program, which is a lot of work and typically done by tenured faculty.

Occasionally I feel overwhelmed by having many services, but I also feel that I am making a difference in the department and that is rewarding.
I am also building a good relationship with my colleagues (and also head-butt with some), and that is important for my tenure case. But of course I don't do this because of my tenure case, I do it because I enjoy it and believe I can make a difference.
In addition, I learn a lot from being in various roles (e.g., how class scheduling work, how the department is run, how to recruit students and recruit faculty, how to deal with difficult people and situation, etc).

In general, I was way more active than most junior faculty in the department.  I was also a lot more involved in the department than I was in my previous department at UNL. Here I share the vision of the department and want to help achieve it.
In short, if you want to make a difference, and \emph{feel comfortable} at handling additional load, you can volunteer for more services. Otherwise, you can just ``stay low'',
do the required services and focus on your research and teaching. This is probably the best strategy for most junior faculty.

\section{Service to the University}
In most cases, these are optional services and assistant professors often do not take any of them.  Common examples include serving the faculty senate, scholarship committee, etc. The university often sends out emails asking for volunteers (so somewhat generic and you can safely ignore), and sometimes people (e.g., your dept chair or dean) will directly ask you to serve (in which case you will want to consider it).

\subsection{Advice and Personal Experience} I only have limited experience with these services. I served in a couple of grants and scholarship committees, in which I review applications for scholarship and grants for students. I also help represent the college in serving as judge of the KimmyDuong foundation scholarship. These are in general time-consuming (e.g., the KimmyDuong scholarship reviews and interview process happen during weekends!), but they are also rewarding as you get to know students and help them. In general, I would recommend doing these services only if you have the time.  However, sometimes they can be rewarding and fun, so you might want to try them out.

\section{Other Types of Services}
\paragraph{Writing Recommendation Letters} You will be asked to write recommendation letters for both students and colleagues.
